Recurrent Deep Learning Models and its applications

### 2025-03-29 paper2/main.tex
fix bug

% Deep recurrent neural networks (DRNNs)
\subsubsection{Deep recurrent neural networks (DRNNs)}
Deep architecture networks with multiple layers that can hierarchically learn complex representations \parencite{bengio2009learning}. By extending of this concept, stacking recurrent layers to form Deep Recurrent Neural Networks aligns with this principle. A number of research papers have been prove that the performance of DRNNs is out-performance then conventional RNNs. \parencite{le2010deep, delalleau2011shallow, pascanu2013on}
\\[1ex]
The architecture of DRNNs are similar with conventional RNNs. We simply stack the recurrent layers vertically. For the first layer, this layer receives the input and combines it with its previous hidden state $h_{t-1}$ (Equation 38). The second layer receive the hidden state of the first layer and treat as the input of the layer 2 (Equation 39). We can extend this concept to $L$ layers (Equation 40).
\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.5\textwidth]{../Pic/drnn.png} % Replace "example-image" with your image file
\end{figure}
\\[1ex]
\begin{equation}
    h^{(1)}_t = f(W^{(1)}_{xh} x_{t} + W^{(1)}_{hh} h^{(1)}_{t-1} + b^{(1)}_h)
\end{equation}
\begin{equation}
    h^{(2)}_t = f(W^{(2)}_{xh} h^{(1)}_{t} + W^{(2)}_{hh} h^{(2)}_{t-1} + b^{(2)}_h)
\end{equation}
\begin{equation}
    h^{(L)}_t = f(W^{(L)}_{xh} h^{(L-1)}_{t} + W^{(L)}_{hh} h^{(L)}_{t-1} + b^{(L)}_h)
\end{equation}
\begin{equation}
    o_t = W_{hy}h^{(L)}_{t} + b_y
\end{equation}
\begin{equation}
    y_t = g(o_t)
\end{equation}
Where $x_t$ is the input at time $t$. $h^{(l)}_{t}$ is the hidden state for the $l$ layer at time t. $W^{(l)}_{xh}, W^{(l)}_{hh}$ are the weight matrices for the input to hidden and hidden to hidden connections in layer $l$, respectively. $W_{hy}$ is weight matrix for the output layer. $b^{(l)}_h$ is the bias vector for the $l$ layer (except output layer), $b_y$ is the bias vector for output layer. $g(\cdot)$ and $f(\cdot)$ are an activation function.

\subsubsection{RNN encoder decoder model}
RNN Encoder-Decoderis a RNN based architecture which commonly used for handle sequence to sequence problem, for example machine translation, summarization, and question answering. The designed of this architecture is good for handle the input sequence and the output sequence is not the same length, leveraging the sequential nature of data. 
\\[1ex]
In this architecture, there are two main components, the first component is rnn encoder, the usage of the encoder is to encode the input sequence into a fixed length vector. And the decoder is to decode the vector into the target sequence. 
\\[1ex]
\textbf{Encoder}
\\[1ex]
The encoder is a RNN based block, which processes the input sequence step by step and calculate the hidden state representation for each input token. Let the input sequence be $X = [x_1, x_2, \cdots, x_T]$ where $T$ is the length of the input sequence, and $x_t$ is the input at time step $t$. We use the below formula to computes the hidden states $h_t$.
\\[1ex]
For the Vanilla RNN encoder
\begin{equation}
    h_t = f_{enc}(x_t, h_{t-1})
\end{equation}

@@

>>>>

% Deep recurrent neural networks (DRNNs)
\subsubsection{Deep recurrent neural networks (DRNNs)}
Deep architecture networks with multiple layers that can hierarchically learn complex representations \parencite{bengio2009learning}. By extending of this concept, stacking recurrent layers to form Deep Recurrent Neural Networks aligns with this principle. A number of research papers have been prove that the performance of DRNNs is out-performance then conventional RNNs. \parencite{le2010deep, delalleau2011shallow, pascanu2013on}
\\[1ex]
The architecture of DRNNs are similar with conventional RNNs. We simply stack the recurrent layers vertically. For the first layer, this layer receives the input and combines it with its previous hidden state $h_{t-1}$ (Equation 38). The second layer receive the hidden state of the first layer and treat as the input of the layer 2 (Equation 39). We can extend this concept to $L$ layers (Equation 40).
\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.5\textwidth]{../Pic/drnn.png} % Replace "example-image" with your image file
\end{figure}
\\[1ex]
\begin{equation}
    h^{(1)}_t = f(W^{(1)}_{xh} x_{t} + W^{(1)}_{hh} h^{(1)}_{t-1} + b^{(1)}_h)
\end{equation}
\begin{equation}
    h^{(2)}_t = f(W^{(2)}_{xh} h^{(1)}_{t} + W^{(2)}_{hh} h^{(2)}_{t-1} + b^{(2)}_h)
\end{equation}
\begin{equation}
    h^{(L)}_t = f(W^{(L)}_{xh} h^{(L-1)}_{t} + W^{(L)}_{hh} h^{(L)}_{t-1} + b^{(L)}_h)
\end{equation}
\begin{equation}
    o_t = W_{hy}h^{(L)}_{t} + b_y
\end{equation}
\begin{equation}
    y_t = g(o_t)
\end{equation}
Where $x_t$ is the input at time $t$. $h^{(l)}_{t}$ is the hidden state for the $l$ layer at time t. $W^{(l)}_{xh}, W^{(l)}_{hh}$ are the weight matrices for the input to hidden and hidden to hidden connections in layer $l$, respectively. $W_{hy}$ is weight matrix for the output layer. $b^{(l)}_h$ is the bias vector for the $l$ layer (except output layer), $b_y$ is the bias vector for output layer. $g(\cdot)$ and $f(\cdot)$ are an activation function. 
\newpage

\subsubsection{RNN encoder decoder model}
\begin{figure}[!htb]
    \centering
    \includegraphics[width=1\textwidth]{../Pic/RNN_encoder_decoder.jpg} % Replace "example-image" with your image file
\end{figure}
RNN Encoder-Decoderis a RNN based architecture which commonly used for handle sequence to sequence problem, for example machine translation, summarization, and question answering. The designed of this architecture is good for handle the input sequence and the output sequence is not the same length, leveraging the sequential nature of data. 
In this architecture, there are two main components, the first component is rnn encoder, the usage of the encoder is to encode the input sequence into a fixed length vector. And the decoder is to decode the vector into the target sequence. 
\\[1ex]
\textbf{Encoder}
\\[1ex]
The encoder is a RNN based block, which processes the input sequence step by step and calculate the hidden state representation for each input token. Let the input sequence be $X = [x_1, x_2, \cdots, x_T]$ where $T$ is the length of the input sequence, and $x_t$ is the input at time step $t$. We use the below formula to computes the hidden states $h_t$.
\begin{equation}
    h_t = f_{enc}(x_t, h_{t-1})
\end{equation}

@@

\subsubsection{Attention Mechanism}
Vaswani et al. introduces the Transformer architecture, which replaces RNN model with self attention. In this section, I want to review the attention machanism. \parencite{vaswani2023attentionneed}. Attention machanism is good for handle the sequence-to-sequence tasks such as machine translation, text summarization and image captioning. Bahdanau et al. \parencite{bahdanau2016neuralmachinetranslationjointly} introduces the attention mechanism in the machine translation task. At each decoding step $t'$, the model model will compute the attention scores $\alpha_{t',t}$ that represent how important of the current hidden state $h_t$ to the current decoder state $s_{t'}$. We have different attention formula. 
\\[1ex]
\textbf{Bahdanau (Additive) Attention}
\\[1ex]
This method proposed by \parencite{bahdanau2016neuralmachinetranslationjointly}, in which this attention uses a feedforward network to compute the attention scores. 
\\[1ex]
\begin{equation}
    e_{t',t} = v_a^Ttanh(W_as_{t'-1}+U_ah_t)
\end{equation}                    
\textbf{Luong (Multiplicative) Attention}
\\[1ex]
This method is introduced by Luong et al \parencite{luong2015effectiveapproachesattentionbasedneural}. This method uses the dot product to compute the attention scores.
\begin{equation}
    e_{t',t} = s_{t'-1}^Th_t
\end{equation}
\textbf{Scaled Dot-Product Attention}
\\[1ex]
